---
title: "Week 10 Homework"
author: "Anna Hiltner Nouzeilles"
format: 
  html:
    embed-resources: true
editor: visual
---

```{r}
#| echo: false
#| message: false
#| warning: false

library(gssr)
library(tidyverse)

theme_set(theme_light())

gss2022 <- gss_get_yr(year = 2022) |> 
  haven::zap_labels() 
```

# Marijuana Legalization and Sex

## Contingency Table
```{r}
#| echo: false
#| message: false
#| warning: false

sgrass <- gss2022 |>
  select(grass, sex) |>
  drop_na() |>
  mutate(
    grass = if_else(grass == 1, 1, 0),
    sex = if_else(sex == 1, 1, 0)
  ) |>
  select(grass, sex)
  
sgrass_table <- table(sgrass$sex, sgrass$grass)

dimnames(sgrass_table) <- list(
  "Sex" = c("Female", "Male"),
  "Support Legalization" = c("Oppose", "Support")
)

print(sgrass_table)

```
## Visualization
```{r}
#| echo: false
#| message: false
#| warning: false

sgrass_bar <- gss2022 |>
  select(grass, sex) |>
  drop_na() |>
  mutate(
    grass = if_else(grass == 1, 1, 0)
  ) |>
  group_by(sex) |>
  summarize(grass_support = mean(grass))

ggplot(sgrass_bar, aes(x = factor(sex), y = grass_support * 100)) +
  geom_col(fill = c("darkgreen", "darkblue")) +
  scale_x_discrete(labels = c("2" = "Female", "1" = "Male")) +
  labs(
    title = "Support for Marijuana Legalization by Sex",
    caption = "Source: 2022 GSS",
    x = "Sex",
    y = "% Supporting Legalization",
  ) 
```

grass = Outcome Y

sex = Predictor X

X = 1 is male

X = 0 is female

## Difference in Probability Regression
```{r}
mytidy <- function(x) {
  broom::tidy(x) |> 
    select(term, estimate, std.error)
}

mod1_sgrass <- lm(grass ~ sex,
            data = sgrass)

mytidy(mod1_sgrass)
```
Males are 3.61 percentage points more likely to support marijuana legalization than females. However, the p-value of .188 is above alpha 0.05 which means we fail to reject the null hypothesis.

## Risk Ratio Regression
```{r}
mod2_sgrass <- glm(grass ~ sex,
            data = sgrass,
            family = binomial(link = "log"))

mytidy(mod2_sgrass)

exp(0.05139)
```
The Risk Ratio is 1.0527. Males are 5.3% more likely to support marijuana legalization compared to females. However, the p-value of 0.186 is higher than our alpha of 0.05 which means we fail to reject the null hypothesis. 

## Log Probability and Risk Ratio
```{r}
mod3_sgrass <- glm(grass ~ sex,
            data = sgrass,
            family = binomial(link = "log"))
mytidy(mod3_sgrass)
```

### Log Probability

$\alpha$ = -0.38

### Log Risk Ratio

$\beta$ = 0.051

## Log Odds

```{r}
mod4_sgrass <- glm(grass ~ sex,
            data = sgrass,
            family = binomial(link = "logit"))

mytidy(mod4_sgrass)
```


$$
\log\left(\frac{P(grass = 1)}{1 - P(grass = 1)}\right)
$$

The log odds that grass = 1 when sex = 0 is $\alpha$ = 0.77601

For sex = 1, the log odds that grass = 1 is 0.77601 + 0.17308 = 0.94909

## Log Odds Ratio

$$
\beta = \log \left[ \frac{P(grass = 1 \mid sex=1)}{1 - P(grass = 1 \mid sex = 1)} \right] - \log \left[ \frac{P(grass = 1 \mid sex=0)}{1 - P(grass = 1 \mid sex=0)} \right]
$$

$\beta$ = 0.17308
The exponentiated beta is 1.1889

The odds of supporting marijuana are 1.19 times greater for males than for females.

```{r}
#| echo: false
#| message: false
#| warning: false

exp(0.17308)
```


# Degree Attainment and Abortion

## Contingency Table
```{r}
degree_abany <- gss2022 |> 
  select(abany, degree) |> 
  drop_na() |> 
  mutate(abany = if_else(abany == 1, 1, 0)) |> 
  mutate(grad = if_else(degree == 4, 1, 0)) |>
  select(abany, grad) 

degree_abany_table <- table(degree_abany$abany, degree_abany$grad)

dimnames(degree_abany_table) <- list(
  "Abortion" = c("Oppose", "Support"),
  "Degree Attainment" = c("Non-Graduate", "Graduate")
)

print(degree_abany_table)
```
## Visualization
```{r}
#| echo: false
#| message: false
#| warning: false

degree_abany_graph <- gss2022 |> 
  select(abany, degree) |> 
  drop_na() |> 
  mutate(abany = if_else(abany == 1, 1, 0), 
         grad = if_else(degree == 4, 1, 0)) |> 
  group_by(grad) |> 
  summarize(abortion_support = mean(abany)) 

ggplot(degree_abany_graph, aes(x = factor(grad), y = abortion_support * 100)) +
  geom_col(fill = c("#ff7261", "#057f8d")) +
  scale_x_discrete(labels = c("0" = "No Graduate Degree", "1" = "Graduate Degree")) +
  labs(
    title = "Support for Abortion by Degree Attainment",
    x = "Degree Attainment",
    y = "% Supporting Abortion",
    caption = "Source: 2022 GSS"
  )

```

## Difference in Probability Regression
```{r}

mod1_degree_abany <- lm(abany ~ grad,
            data = degree_abany)

mytidy(mod1_degree_abany)
```
Grads are 13.4% percentage points more likely to support abortion than non-graduates. 

## Risk Ratio Regression
```{r}
mod2_degree_abany <- glm(abany ~ grad,
            data = degree_abany,
            family = binomial(link = "log"))

mytidy(mod2_degree_abany)

exp(0.210283)
```
The Risk Ratio is 1.234. Those with a grad degree are are 23% more likely to support abortion compared to those without a grad degree. 

## Log Probability and Risk Ratio
```{r}
mod3_degree_abany <- glm(abany ~ grad,
            data = degree_abany,
            family = binomial(link = "log"))
mytidy(mod3_degree_abany)
```

### Log Probability

$\alpha$ = -0.555

### Log Risk Ratio

$\beta$ = 0.21

## Log Odds

```{r}
mod4_degree_abany <- glm(abany ~ grad,
            data = degree_abany,
            family = binomial(link = "logit"))

mytidy(mod4_degree_abany)

exp(.59)
```


$$
\log\left(\frac{P(abany = 1)}{1 - P(abany = 1)}\right)
$$

The log odds that abany = 1 when grad = 0 is $\alpha$ = .299

For grad = 1, the log odds that abany = 1 is 0.299 + 0.589 = 0.888

## Log Odds Ratio

$$
\beta = \log \left[ \frac{P(abany = 1 \mid grad=1)}{1 - P(abany = 1 \mid grad = 1)} \right] - \log \left[ \frac{P(abany = 1 \mid grad=0)}{1 - P(abany = 1 \mid grad=0)} \right]
$$

$\beta$ = 0.59
The exponentiated beta is 1.8

The odds of supporting abortion are 1.8 times greater for those with a graduate degree than those without one.

# Voted for Biden and Marijuana Legalization

## Contingency Table
```{r}
#| echo: false
#| message: false
#| warning: false

biden_grass <- gss2022 |> 
  select(pres20, grass) |>
  drop_na() |>
  mutate(biden = if_else(pres20 == 1, 1, 0),
         grass = if_else(grass == 1, 1, 0)) |>
  select(biden, grass) 

biden_grass_table <- table(biden_grass$biden, biden_grass$grass)

dimnames(biden_grass_table) <- list(
  "Biden" = c("No Vote", "Vote"),
  "Marijuana Legalization" = c("Oppose", "Support")
)

print(biden_grass_table)
```

## Visualization
```{r}
#| echo: false
#| message: false
#| warning: false

biden_grass_bar <- gss2022 |> 
  select(pres20, grass) |>
  drop_na() |>
  mutate(
    biden = if_else(pres20 == 1, 1, 0),
    grass = if_else(grass == 1, 1, 0)
  ) |>
  group_by(grass) |>
  summarize(vote_biden = mean(biden))

ggplot(biden_grass_bar, aes(x = factor(grass), y = vote_biden * 100)) +
  geom_col(fill = c("darkred", "#057f8d")) +
  scale_x_discrete(labels = c("0" = "Oppose Legalization", "1" = "Support Legalization")) +
  labs(
    title = "Support for Biden by Support of Marijuana Legalization",
    x = "Support of Marijuana Legalization",
    y = "% Voted for Biden",
    caption = "Source: 2022 GSS"
  ) 

```


## Difference in Probability Regression
```{r}
mod1_biden_grass <- lm(biden ~ grass,
            data = biden_grass)

mytidy(mod1_biden_grass)
```
Marijuana legalization supporters are 29 percentage points more likely to have voted for Biden than those opposed to marijuana legalization. 

## Risk Ratio Regression
```{r}
mod2_biden_grass <- glm(biden ~ grass,
            data = biden_grass,
            family = binomial(link = "log"))

mytidy(mod2_biden_grass)

exp(0.5932218)
```
The Risk Ratio is 1.81. Marijuana legalization supporters are 81% more likely to have voted for Biden than those who oppose marijuana legalization. 

## Log Probability and Risk Ratio
```{r}
mod3_biden_grass <- glm(biden ~ grass,
            data = biden_grass,
            family = binomial(link = "log"))
mytidy(mod3_biden_grass)
```

### Log Probability

$\alpha$ = -1.03

### Log Risk Ratio

$\beta$ = 0.593

## Log Odds

```{r}
mod4_biden_grass <- glm(biden ~ grass,
            data = biden_grass,
            family = binomial(link = "logit"))

mytidy(mod4_biden_grass)
```


$$
\log\left(\frac{P(biden = 1)}{1 - P(biden = 1)}\right)
$$

The log odds that sex = 1 when grass = 0 is $\alpha$ = -0.58

For grass = 1, the log odds that sex = 1 is -.582 + 1.2 = 0.618

## Log Odds Ratio

$$
\beta = \log \left[ \frac{P(sex = 1 \mid grass=1)}{1 - P(sex = 1 \mid grass = 1)} \right] - \log \left[ \frac{P(sex = 1 \mid grass=0)}{1 - P(sex = 1 \mid grass=0)} \right]
$$

$\beta$ = 1.2
The exponentiated beta is 3.32

The odds of being a male are 3.32 times greater for marijuana legalization supporters than for those opposed.

```{r}
#| echo: false
#| message: false
#| warning: false

exp(1.2)
```


# Voted for Biden and Degree Attainment

## Contingency Table
```{r}
#| echo: false
#| message: false
#| warning: false

biden_grad <- gss2022 |> 
  select(pres20, degree) |>
  drop_na() |>
  mutate(biden = if_else(pres20 == 1, 1, 0),
         grad = if_else(degree == 4, 1, 0)) |>
  select(biden, grad) 

biden_grad_table <- table(biden_grad$biden, biden_grad$grad)

dimnames(biden_grad_table) <- list(
  "Biden" = c("No Vote", "Vote"),
  "Degree Attainment" = c("Non-Graduate", "Graduate")
)

print(biden_grad_table)
```

## Visualization
```{r}
#| echo: false
#| message: false
#| warning: false

biden_grad_bar <- gss2022 |> 
  select(pres20, degree) |> 
  drop_na() |> 
  mutate(
    biden = if_else(pres20 == 1, 1, 0),
    grad = if_else(degree == 4, 1, 0)
  ) |>
  group_by(grad) |> 
  summarize(vote_biden = mean(biden))

ggplot(biden_grad_bar, aes(x = factor(grad), y = vote_biden * 100)) +
  geom_col(fill = c("orange", "purple")) +
  scale_x_discrete(labels = c("0" = "No Graduate Degree", "1" = "Graduate Degree")) +
  labs(
    title = "Support for Biden by Degree Attainment",
    x = "Degree Attainment",
    y = "% Voted for Biden",
    caption = "Source: 2022 GSS"
  ) 
```

## Difference in Probability Regression
```{r}
mod1_biden_grad <- lm(biden ~ grad,
            data = biden_grad)

mytidy(mod1_biden_grad)
```
Respondents with a graduate degree are 17.2 percentage points more likely to have voted for Biden than those without a graduate degree.

## Risk Ratio Regression
```{r}
mod2_biden_grad <- glm(biden ~ grad,
            data = biden_grad,
            family = binomial(link = "log"))

mytidy(mod2_biden_grad)

exp(0.2566021)
```
The Risk Ratio is 1.29. Those with a graduate degree are 29.2% more likely to have voted for Biden than those who do not have a graduate degree. 

## Log Probability and Risk Ratio
```{r}
mod3_biden_grad <- glm(biden ~ grad,
            data = biden_grad,
            family = binomial(link = "log"))
mytidy(mod3_biden_grad)
```

### Log Probability

$\alpha$ = -0.53

### Log Risk Ratio

$\beta$ = 0.26

## Log Odds

```{r}
mod4_biden_grad <- glm(biden ~ grad,
            data = biden_grad,
            family = binomial(link = "logit"))

mytidy(mod4_biden_grad)
```


$$
\log\left(\frac{P(biden = 1)}{1 - P(biden = 1)}\right)
$$

The log odds that grass = 1 when grad = 0 is $\alpha$ = 0.359

For grad = 1, the log odds that grass = 1 is 0.36 + 0.8 = 1.16

## Log Odds Ratio

$$
\beta = \log \left[ \frac{P(grass = 1 \mid grad=1)}{1 - P(grass = 1 \mid grad = 1)} \right] - \log \left[ \frac{P(grass = 1 \mid grad=0)}{1 - P(grass = 1 \mid grad=0)} \right]
$$

$\beta$ = .799
The exponentiated beta is 2.22

The odds of supporting marijuana legalization are 2.22 times greater for those with a grad degree than those without.

```{r}
#| echo: false
#| message: false
#| warning: false

exp(.799)
```

