---
title: "Week 12 Homework"
author: "Anna Hiltner Nouzeilles"
format: 
  html:
    embed-resources: true
editor: visual
---

```{r}
#| echo: false
#| message: false
#| warning: false

library(gssr)
library(tidyverse)
library(broom)
library(here)

gss <- gss_get_yr(year = 2022) |> 
  haven::zap_labels() 
```

## Linear Regression (LPM) #1

$$
\text{E}(\text{proweed}) = \alpha + \beta_1 \cdot \text{relaxed} + \beta_2 \cdot \text{college} + \beta_3 \cdot (\text{relaxed} \cdot \text{college})
$$

```{r}
#| echo: false
#| message: false
#| warning: false


d <- gss |>
  select(
    grass,
    hrsrelax,
    degree) |>
  drop_na() |>
  mutate(
    proweed = if_else(grass == 1, 1, 0),
    relaxed = if_else(hrsrelax >= 3, 1, 0),
    college = if_else(degree >= 3, 1, 0)
  )

#table of probs

dt <- d |>
  summarize(
    n = n(),
    pgrass = mean(proweed),
    .by = c(relaxed, college)) |>
  arrange(relaxed, college)

print(dt)
```

### Manual Estimation

$\alpha$ = 0.67 which means people who relax less and don't have a college degree have a 67% probability of supporting marijuana legalization.

$\alpha$ + $\beta_1$ = 0.813 therefore b1 = 0.813 - 0.67 = 0.143 which means people who relax more and do not have a college degree are 14.3 percentage points more likely to support marijuana legalization than those without a degree.

$\alpha$ + $\beta_2$ = 0.729 therefore b2 = 0.729 - 0.67 = 0.059 which means those who relax less and have a college degree are 5.9 percentage points more likely to support marijuana legalization.

$\alpha$ + $\beta_1$ + $\beta_2$ + $\beta_3$ = 0.787 therefore b3 = 0.787 - 0.872 = -0.085 which means a person who relaxes more and has a college degree is 8.5 percentage points more likely to support marijuana than a person who relaxes less and does not have a college degree.

### Confirm

```{r}
lpm <- lm(proweed ~ college * relaxed,
          data = d)

tidy(lpm) |> select(term, estimate)
```

### Restricted Model

```{r}
lpm3 <- lm(proweed ~ college + relaxed,
          data = d)

tidy(lpm3) |> select(term, estimate)

AIC(lpm) 
AIC(lpm3)
```

The restricted model is preferred because the AIC is lower.

```{r}
#| echo: false
#| message: false
#| warning: false

# Can also do Likelihood Ratio Test

lrt3 <- anova(lpm3, lpm, test = "Chisq")

tidy(lrt3)
```

## Logit #1

```{r}
dt <- dt |>
  mutate(logodds = log(pgrass / (1 - pgrass)))

dt
```

### Manual Estimation

$\alpha$ = 0.71 which means among people who relax 2 or less hours a day and do not have a college degree, the log odds of supporting marijuana legalization is 0.71

$\alpha$ + $\beta_1$ = 1.47 therefore $\beta_1$ = 1.47 - 0.71 = 0.76 which means among non-college grads, being more relaxed decreases the log odds of supporting marijuana legalization by 0.76.

$\alpha$ + $\beta_2$ = 0.99 therefore $\beta_2$ = 0.99 - 0.71 = 0.28 which means among those who relax less, having a college degree increases the log odds of supporting marijuana legalization by 0.28.

$\alpha$ + $\beta_1$ + $\beta_2$ + $\beta_3$ = 1.31 therefore $\beta_3$ = 1.31 - 1.75 = -0.44 which means the interaction between relaxing more and having a college degree adds a 0.44 to the log odds of supporting marijuana legalization.

### Confirm

```{r}
logitm <- glm(proweed ~ college * relaxed,
              data = d,
              family = binomial(link = "logit"))

tidy(logitm) |> select(term, estimate)
```

### Restricted Model

```{r}
logitm1 <- glm(proweed ~ college + relaxed,
              data = d,
              family = binomial(link = "logit"))

tidy(logitm1) |> select(term, estimate)

AIC(logitm) 
AIC(logitm1)
```

```{r}
#| echo: false
#| message: false
#| warning: false

# Can also do the Likelihood Ratio Test

lrt4 <- anova(logitm1, logitm, test = "Chisq")

tidy(lrt4)
```

The restricted model is preferred because the AIC is lower.

## Poisson #1

```{r}
#| echo: false
#| message: false
#| warning: false

d1 <- gss |>
  select(
    nmbrkids,
    hrsrelax,
    degree) |>
  drop_na() |>
  mutate(
    relaxed = if_else(hrsrelax >= 3, 1, 0),
    college = if_else(degree >= 3, 1, 0),
  )

#table of probs

dt1 <- d1 |>
  summarize(
    n = n(),
    pkids = mean(nmbrkids),
    .by = c(relaxed, college)) |>
  arrange(relaxed, college)

#or

poisson_dt <- d1 |>
  group_by(relaxed, college) |> 
  summarize(
    n = n(),
    mean_kids = mean(nmbrkids))

print(poisson_dt)
```

$\alpha$ = log(2.285) which means people who relax less and do not have a college degree have an estimated 0.83 log children.

$\alpha$ + $\beta_1$ = log(1.857) therefore $\beta_1$ = log(1.857) - 0.83 = -0.211 which means a person who does not have a college degree but relaxes more has 0.211 less log children.

$\alpha$ + $\beta_2$ = log(1.932) therefore $\beta_2$ = log(1.932) - 0.83 = -0.17 which means among those who relax less, having a college degree decreases a person's number of log children by 0.17 log siblings

$\alpha$ + $\beta_1$ + $\beta_2$ + $\beta_3$ = log(1.64) therefore $\beta_3$ = log(1.64) - 0.449 = 0.046 which means more relaxed, college-educated people have 0.046 more log children than less relaxed, non-college educated people.

### Confirm

```{r}
poisson_d <- glm(nmbrkids ~ college * relaxed,
                data = d1,
                family = poisson(link = "log"))

tidy(poisson_d) |> select(term, estimate)
```

### Restricted Model

```{r}
poisson_d1 <- glm(nmbrkids ~ college + relaxed,
                data = d1,
                family = poisson(link = "log"))

tidy(poisson_d1) |> select(term, estimate)

lrt5 <- anova(poisson_d, poisson_d1, test = "Chisq")

AIC(poisson_d)
AIC(poisson_d1)
```

The restricted model is preferred here because its AIC is lower.

## Linear Regression (LPM) #2

$$
\text{E}(\text{liberal}) = \alpha + \beta_1 \cdot \text{relaxed} + \beta_2 \cdot \text{college} + \beta_3 \cdot (\text{relaxed} \cdot \text{college})
$$

```{r}
#| echo: false
#| message: false
#| warning: false

k <- gss |>
  select(polviews,
         hrsrelax,
         degree) |>
  drop_na() |>
  mutate(liberal = if_else(polviews < 4, 1, 0),
         relaxed = if_else(hrsrelax >= 3, 1, 0),
         college = if_else(degree >= 3, 1, 0))

kt <- k |>
  summarize(n = n(),
            plib = mean(liberal),
            .by = c(relaxed, college)) |>
  arrange(relaxed, college)

print(kt)
```

### Manual

$\alpha$ = 0.253 which means among people without a college degree who relax 2 or less hours per day, the probability of being liberal is 25.3%

$\alpha$ + $\beta_1$ = 0.252 therefore $\beta_1$ = 0.252 - 0.253 = -0.001 which means among people without a college degree, those who relax 3 or more hours a day are 0.1 percentage points less likely to be liberal than those who relax less.

$\alpha$ + $\beta_2$ = 0.395 therefore $\beta_2$ = 0.395 - 0.253 = 0.142 which means among people who relax 2 or less hours a day, those with a college degree are 14.2 percentage points more likely to be liberal than those without.

$\alpha$ + $\beta_1$ + $\beta_2$ + $\beta_3$ = 0.47 therefore $\beta_3$ = 0.47 - 0.394 = 0.076 which means the additional effect of being both relaxed and college educated on the probability of being liberal is an increase of 7.6 percentage points.

### Confirm

```{r}
lpm1 <- lm(liberal ~ college * relaxed,
          data = k)

tidy(lpm) |> select(term, estimate)
```

### Restricted Model

```{r}
lpm2 <- lm(liberal ~ college + relaxed,
          data = k)

tidy(lpm2) |> select(term, estimate)
```

```{r}
#| echo: false
#| message: false
#| warning: false

#Likelihood Ratio Test

lrt <- anova(lpm2, lpm1, test = "Chisq")

tidy(lrt)
```

```{r}
AIC(lpm1)
AIC(lpm2)
```

Since a lower AIC is better, the saturated model is preferred.

## Logit #2

### Manual

```{r}
#add log odds
kt <- kt |>
  mutate(logodds = log(plib / (1 - plib)))

kt

```

$\alpha$ = -1.08 which means among people who relax 2 or less hours a day and do not have a college degree, the log odds of being liberal is -1.08.

$\alpha$ + $\beta_1$ = -1.09 therefore $\beta_1$ = -1.09 + 1.08 = -0.01 which means among non-college grads, being more relaxed decreases the log odds of being liberal by 0.01, or essentially no effect.

$\alpha$ + $\beta_2$ = -0.43 therefore $\beta_2$ = -0.43 + 1.08 = 0.65 which means among those who relax less, having a college degree increases the log odds of being liberal by 0.65, or a 1.91x increase in the odds of being liberal.

$\alpha$ + $\beta_1$ + $\beta_2$ + $\beta_3$ = -0.14 therefore $\beta_3$ = -0.14 + 0.44 = 0.3 which means the interaction between relaxing more and having a college degree adds an 0.3 to the log odds of being liberal.

### Confirm

```{r}
logitk <- glm(liberal ~ college * relaxed,
              data = k,
              family = binomial(link = "logit"))

tidy(logitk) |> select(term, estimate)
```

### Restricted Model

```{r}
logitk1 <- glm(liberal ~ college + relaxed,
              data = k,
              family = binomial(link = "logit"))

tidy(logitk1) |> select(term, estimate)

AIC(logitk) 
AIC(logitk1)
```

The saturated model is preferred as its AIC is lower.

```{r}
#| echo: false
#| message: false
#| warning: false

#Can also do Likelihood Ratio Test

lrt1 <- anova(logitk1, logitk, test = "Chisq")

tidy(lrt1)
```

## Poisson #2

\log(\lambda)

$$
\log(\lambda) = \alpha + \beta_1 \cdot \text{relaxed} + \beta_2 \cdot \text{college} + \beta_3 \cdot (\text{relaxed} \cdot \text{college})
$$

### Manual

```{r}
#| echo: false
#| message: false
#| warning: false

k1 <- gss |>
  select(
    sibs,
    hrsrelax,
    degree) |>
  drop_na() |>
  mutate(
    relaxed = if_else(hrsrelax >= 3, 1, 0),
    college = if_else(degree >= 3, 1, 0),
  )

#table of probs

kt1 <- k1 |>
  summarize(
    n = n(),
    psibs = mean(sibs),
    .by = c(relaxed, college)) |>
  arrange(relaxed, college)

#or

poisson_kt <- k1 |>
  group_by(relaxed, college) |> 
  summarize(
    n = n(),
    mean_sibs = mean(sibs))

poisson_kt
```

$\alpha$ = log(3.84) which means people who relax less and do not have a college degree have an estimated 1.35 log siblings.

$\alpha$ + $\beta_1$ = log(3.55) therefore $\beta_1$ = log(3.55) - 1.35 = -0.08 which means people who relax more and do not have a college degree have 0.08 less log siblings

$\alpha$ + $\beta_2$ = log(2.43) therefore $\beta_2$ = log(2.43) -1.35 = -0.46 which means a person who has a college degree but relaxes less has 0.46 less log siblings.

$\alpha$ + $\beta_1$ + $\beta_2$ + $\beta_3$ = log(2.52) therefore $\beta_3$ = 0.11 which means more relaxed, college-educated people have 0.11 more log siblings than less relaxed, non-college educated people.

### Confirm

```{r}
poisson_k1 <- glm(sibs ~ college * relaxed,
                data = k1,
                family = poisson(link = "log"))

tidy(poisson_k1) |> select(term, estimate)
```

### Restricted Model

```{r}
poisson_k2 <- glm(sibs ~ college + relaxed,
                data = k1,
                family = poisson(link = "log"))

tidy(poisson_k2) |> select(term, estimate)

AIC(poisson_k1) 
AIC(poisson_k2)
```

The saturated model is preferred as its AIC is much lower.

```{r}
#| echo: false
#| message: false
#| warning: false

#Can also do Likelihood Ratio Test

lrt2 <- anova(poisson_k1, poisson_k2, test = "Chisq")

tidy(lrt2)
```
