---
title: "Week 11 Homework"
author: "Anna Hiltner Nouzeilles"
format: 
  html:
    embed-resources: true
editor: visual
---

```{r}
#| echo: false
#| message: false
#| warning: false

library(gssr)
library(tidyverse)

theme_set(theme_light())

gss2022 <- gss_get_yr(year = 2022) |> 
  haven::zap_labels() 
```

# Gun Control and Sex

## LPM (or Difference in Probability Regression)

```{r}
#| echo: false
#| message: false
#| warning: false

sgun <- gss2022 |> 
  select(gunlaw, sex) |> 
  drop_na() |> 
  mutate(
    gunlaw = if_else(gunlaw == 1, 1, 0),
    sex = if_else(male == 1, 1, 0)
  ) |>
  select(gunlaw, male)

sgun_table <- table(sgun$male, sgun$gunlaw)

dimnames(sgun_table) <- list(
  "Sex" = c("Female", "Male"),  
  "Gun Control" = c("Oppose", "Support") 
)

print(sgun_table)
```

**Outcome:** Gun Control (Support/Oppose)

**Predictor:** Sex (Male/Female)

### First, Manually:

Female conditional probability of support: 1143/1431 = 0.7987

Male conditional probability of support: 816/1265 = 0.6451

0.7987 - 0.6451 = .1536 which means that females are 15.4% more likely to support gun control than males.

### In R:

```{r}
mytidy <- function(x) {
  broom::tidy(x) |> 
    select(term, estimate, std.error)
}

mod1_sgun <- lm(gunlaw ~ male,
            data = sgun,
            family = binomial(link = "identity"))

mytidy(mod1_sgun)
```

The intercept is the predicted probability of supporting gun control for females. The sex coefficient is the difference in probability between males and females.

Men are 15.4 percentage points less likely to support gun control.

## Risk Ratio

### First, Manually:

Male Probability of Support / Female Probability of Support = .6451 / .7987 = 0.8076

### In R:

```{r}
mod2_sgun <- glm(gunlaw ~ male,
            data = sgun,
            family = binomial(link = "log"))

mytidy(mod2_sgun)

exp(-.21370)
```

The risk ratio of .81 means that men are 81% as likely as women to support gun control, or (1-.81=) 19% less likely to support gun control than women.

## Logit

```{r}
mod3_sgun <- glm(gunlaw ~ male,
            data = sgun,
            family = binomial(link = "logit"))

mytidy(mod3_sgun)

exp(-.7810)
```

The log odds that females support gun control (gunlaw = 1 when sex = 0) is $\alpha$ = 1.38

The log odds that males support gun control (sex = 1 when gunlaw = 1) is 1.38 - 0.7810 = .599

The odds ratio is exp(-0.781) = 0.4577 which means that the odds of men supporting gun control are 46% of the odds for women.

## Log Likelihood Comparison

```{r}
#using binomial distribution

sgun <- sgun |>
  mutate(predprob = 
           predict(mod1_sgun,
                   type = "response")) |>
  mutate(l = dbinom(gunlaw, 1, predprob),
         ll = log(l))

sum(sgun$ll)
logLik(mod1_sgun)
#Is it an issue that the results are slightly different? Why are the other log likelihood comparisons exactly the same while this one isn't?
```

# Linear Regressions (Gaussian GLMs)

### First Gaussian GLM

```{r}
sexweight <- gss2022 |>
  select(weight, sex) |>
  drop_na() |>
  mutate(male = if_else(sex <= 1, 1, 0)) |>
  select(weight, male)

mod1_sexweight <- lm(weight ~ male,
                     data = sexweight)

mytidy(mod1_sexweight)
```

The average weight for females is 173 lbs (alpha). Males weigh 24 lbs (beta) more than females. (173 + 24 = 197) The average weight for males is 197 lbs.

#### Log Likelihood Comparison

```{r}
sexweight <- sexweight |>
  mutate(mean = predict(mod1_sexweight),
         sd = sd(mod1_sexweight$residuals)) |>
  mutate(l = dnorm(weight, mean = mean, sd = sd),
         ll = dnorm(weight, mean = mean, sd = sd, 
                    log = TRUE))

sum(sexweight$ll)
logLik(mod1_sexweight)
```

### Second Gaussian GLM

```{r}
sextvhours <- gss2022 |>
  select(tvhours, sex) |>
  drop_na() |>
  mutate(male = if_else(sex <= 1, 1, 0)) |>
  select(tvhours, male)

mod1_sextvhours <- lm(tvhours ~ male,
                     data = sextvhours)

mytidy(mod1_sextvhours)
```

Females watch 3.44 (alpha) hours of TV on average. Males watch .065 (beta) less hours of TV than females, which means they watch 3.375 hours of TV on average.

# Poisson GLM

```{r}
sexpartners <- gss2022 |>
  select(partnrs5, sex) |>
  drop_na() |>
  mutate(male = if_else(sex <= 1, 1, 0)) |>
  select(partnrs5, male)

mod1_sexpartners <- glm(partnrs5 ~ male,
               data = sexpartners,
               family = poisson(link="log"))

sexpartners_table <- table(sexpartners$male, sexpartners$partnrs5)

print(sexpartners_table)
mytidy(mod1_sexpartners)

exp(0.3703)
exp(0.2155)
exp(0.59)
```

Alpha is .37 log partners for women. Exponentiated, we find that women had on average 1.45 partners in the last 5 years.

.37 + .216 = .59 log partners for men. Exponentiated, we find that men had on average 1.8 partners in the last five years.

Beta is 0.216 log partners, which means that men had .216 more log partners than women in the past five years. Exponentiated, there is a ratio of 1.24, which means that men had 24% more partners in the last year than women.

### Manually

```{r}
#| echo: false

gss2022 |>
  select(partnrs5, sex) |>
  drop_na() |>
  mutate(sex = if_else(sex == 1, "male", "female")) |>
  summarize(mean_partners = mean(partnrs5), .by = sex) 
```

$$
  \log(y_i) = \alpha + \beta x_i
$$

log(1.45) = .37 for females

log(1.8) = .59 for males

beta = .59 - .37 = .216 = log rate ratio, or the difference in log sex partners between females and males.

exp(.216) = 1.24, which means men have 1.24x, or 24% more sexual partners than women

In LPM style, 1.8 - 1.45 = .35, so males have .35 more siblings than women.

### Second Poisson GLM

```{r}
sibsdegree <- gss2022 |>
  select(sibs, degree) |>
  drop_na() |>
  mutate(college = if_else(degree <= 3, 1, 0)) |>
  select(sibs, college)

mod1_sibsdegree <- glm(sibs ~ college,
               data = sibsdegree,
               family = poisson(link="log"))

sibsdegree_table <- table(sibsdegree$sibs, sibsdegree$college)

print(sibsdegree_table)
mytidy(mod1_sibsdegree)

exp(0.9525)
exp(0.3130)
exp(0.6395)
```

Alpha is 0.9525 log siblings for those with a college degree. Exponentiated, we find that those with a college degree have on average 2.59 siblings.

0.9525 - 0.3130 = 0.6395 log siblings for those without a college degree. Exponentiated, we find that those without a college degree have on average 1.89 siblings.

Beta is 0.3130 log siblings, which means those with a college degree have 0.313 more siblings than those without a college degree. Exponentiated, we get a ratio of 1.37, which means that those with a college degree have 37% more siblings than those without a college degree.

#### Log Likelihood Comparison

```{r}
sibsdegree <-
  sibsdegree |>
  mutate(mean = predict(mod1_sibsdegree, type = "response")) |>
  mutate(l = dpois(sibs, lambda = mean),
         ll = dpois(sibs, lambda = mean, 
                    log = TRUE))

sum(sibsdegree$ll)
logLik(mod1_sibsdegree)
```
