---
title: "Week 6 Homework"
author: "Anna Hiltner Nouzeilles"
format: 
  html:
    embed-resources: true
editor: visual
---

```{r}
#| echo: false
#| message: false
#| warning: false

library(ggplot2)
library(dplyr)
library(haven)
library(gssr)
library(tidyverse)

theme_set(theme_light())

gss2022 <- gss_get_yr(year = 2022) |> 
  haven::zap_labels() 
```

# Hypothesis Test: GRASS

**Question** : *Do you think the use of marijuana should be made legal or not?*

```{r}
#| echo: false
#| message: false
#| warning: false

g <- gss2022 |>
  select(grass) |>
  drop_na() |>
  mutate(grass = ifelse(grass == 1, 1, 0))

grass_null <- .7

grass_prop <- mean(g$grass)
grass_sd <- sd(g$grass)
grass_n <- nrow(g)
grass_se <- grass_sd / sqrt(grass_n)
grass_diff <- grass_prop - grass_null
grass_zscore <- grass_diff / grass_se
grass_pvalue <- ((1 - pnorm(0.1239)) * 2)

grass_95low <- mean(g$grass) - (1.96 * .01366)
grass_95high <- mean(g$grass) + (1.96 * .01366)
```

I chose a null of **.7**, which assumes that **70%** of the population favors the death penalty

I will use the conventional alpha level of **.05**, which means I am willing to falsely reject the null hypothesis 5% of the time.

The proportion of grass is $\bar{x}$ = **`r round(mean(g$grass, na.rm = TRUE), 2)`**.

The standard deviation of grass is **`r round(sd(g$grass), 2)`**

The standard error of grass is **`r round(grass_sd / sqrt(grass_n), 4)`**

The z-score is **`r round(grass_diff / grass_se, 4)`**

The p-value, or probability of getting an absolute z-score of 0.1239 or greater, is about **`r ((1 - pnorm(0.1239)) * 2)`**

I have a **.9%** probability of getting a sample proportion of at least **.702** if the true population proportion were **.7**. Since the p-value of .9 is higher than alpha level 0.05, we fail to reject the null hypothesis. I can also use the 95% confidence interval for grass of **.675** and **.728** to fail to reject the null hypothesis.

```{r}
#| echo: false
#| message: false
#| warning: false

ggplot(data.frame(x = c(.63, .77)), aes(x = x)) +
  stat_function(fun = dnorm, 
                args = list(mean = grass_null, sd = grass_se), 
                color = "orange") +
  stat_function(fun = dnorm, 
                args = list(mean = grass_null, sd = grass_se), 
                geom = "area", 
                xlim = c(.63, .77), 
                fill = "lightpink", 
                alpha = 0.6) +
  geom_vline(xintercept = grass_prop, 
             linetype = "dashed", 
             color = "black") +
  labs(x = "Proportion Favoring Marijuana Legalization", 
       y = "Density")
```

# Hypothesis Test: The Death Penalty

**Question** : *Do you favor or oppose the death penalty for persons convicted of murder?*

```{r}
#| echo: false
#| message: false
#| warning: false

d <- gss2022 |>
  select(cappun) |>
  drop_na() |>
  mutate(cappun = ifelse(cappun == 1, 1, 0))

dp_prop <- mean(d$cappun)
dp_sd <- sd(d$cappun)
dp_n <- nrow(d)
dp_se <- dp_sd / sqrt(dp_n)
dp_null <- .4
dp_diff <- dp_prop - dp_null
dp_zscore <- dp_diff / dp_se
dp_pvalue <- ((1 - pnorm(28.986)) * 2)

dp_95low <- mean(d$cappun) - (1.96 * .0078)
dp_95high <- mean(d$cappun) + (1.96 * .0078)
```

I chose a null of .4, which assumes that 40% of the population favors the death penalty

I will use the conventional alpha level of .05, which means I am willing to falsely reject the null hypothesis 5% of the time.

The proportion of cappun is $\bar{x}$ = **`r round(mean(d$cappun, na.rm = TRUE), 2)`**.

The standard deviation of cappun is **`r round(sd(d$cappun), 2)`**

The standard error of cappun is **`r round(dp_sd / sqrt(dp_n), 4)`**

The z-score is **`r round(dp_diff / dp_se, 4)`**

The p-value, or probability of getting an absolute z-score of 28.9858 or greater, is about **`r ((1 - pnorm(26.14)) * 2)`**

I have an almost **0%** probability of getting a sample proportion of at least **.604** if the true population proportion were **.4**. Since the p-value of 0 is lower than alpha level 0.05, we can reject the null hypothesis. I can also use the 95% confidence interval for cappun of **.589** and **.619** to reject the null hypothesis.

```{r}
#| echo: false
#| message: false
#| warning: false

ggplot(data.frame(x = c(.35, .65)), aes(x = x)) +
  stat_function(fun = dnorm, 
                args = list(mean = dp_null, sd = dp_se), 
                color = "blue") +
  stat_function(fun = dnorm, 
                args = list(mean = dp_null, sd = dp_se), 
                geom = "area", 
                xlim = c(.35, .65), 
                fill = "violet", 
                alpha = 0.6) +
  geom_vline(xintercept = dp_prop, 
             linetype = "dashed", 
             color = "black") +
  labs(x = "Proportion Favoring Death Penalty", 
       y = "Density")
```

# Hypothesis Test: age

```{r}
#| echo: false
#| message: false
#| warning: false

a <- gss2022 |>
  select(age) |>
  drop_na()

age_xbar <- mean(a$age)
age_sd <- sd(a$age)
age_n <- nrow(a)
age_se <- age_sd / sqrt(age_n)
age_null <- 40
age_diff <- age_xbar - age_null 
age_zscore <- age_diff / age_se
age_pvalue <- ((1 - pnorm(28.986)) * 2)

age_95low <- mean(a$age) - (1.96 * .28)
age_95high <- mean(a$age) + (1.96 * .28)
```

A reasonable null for this question would be **40 years old.**

I will use the conventional alpha level of **.05**, which means I am willing to falsely reject the null hypothesis 5% of the time.

The mean of age is $\bar{x}$ = **`r round(mean(g$age, na.rm = TRUE), 2)`**. This means that, on average, individuals in this sample were moderate.

The standard deviation of age is **`r round(sd(a$age), 2)`**

The standard error of age is **`r round(age_sd / sqrt(age_n), 4)`**

The z-score is **`r round(age_diff / age_se, 4)`**

The p-value, or probability of getting an absolute z-score of 28.9858 or greater, is about **`r ((1 - pnorm(28.9858)) * 2)`**

I have an almost **0%** probability of getting a sample mean of at least **48.244** if the true population mean were **40**. Since the p-value of 0 is lower than alpha level 0.05, we can reject the null hypothesis. I can also use the 95% confidence interval for age of **47.696** and **48.793** to reject the null hypothesis.

```{r}
#| echo: false
#| message: false
#| warning: false

ggplot(data.frame(x = c(38, 50)), aes(x = x)) +
  stat_function(fun = dnorm, 
                args = list(mean = 40, sd = 0.2844), 
                color = "#CC0000") +
  stat_function(fun = dnorm, 
                args = list(mean = 40, sd = 0.2844), 
                geom = "area", 
                xlim = c(39, 41), 
                fill = "lightblue", 
                alpha = 0.6) +
  geom_vline(xintercept = 48.2445, 
             linetype = "dashed", 
             color = "black") +
  labs(x = "Mean Age", 
       y = "Density")
```

# Hypothesis Test: Hours Worked

```{r}
#| echo: false
#| message: false
#| warning: false

h <- gss2022 |>
  select(hrs1) |>
  drop_na()

hrs1_xbar <- mean(h$hrs1)
hrs1_sd <- sd(h$hrs1)
hrs1_n <- nrow(h)
hrs1_se <- hrs1_sd / sqrt(hrs1_n)
hrs1_null <- 40
hrs1_diff <- hrs1_xbar - hrs1_null 
hrs1_zscore <- hrs1_diff / hrs1_se
hrs1_pvalue <- ((1 - pnorm(0.20916)) * 2)

hrs1_95low <- mean(h$hrs1) - (1.96 * .29)
hrs1_95high <- mean(h$hrs1) + (1.96 * .29)
```

A reasonable null for this question would be **40 hours** worked per week.

I will use the conventional alpha level of **.06**, which means I am willing to falsely reject the null hypothesis 6% of the time.

The mean of hrs1 is $\bar{x}$ = **`r round(mean(h$hrs1, na.rm = TRUE), 2)`**.

The standard deviation of hours worked is **`r round(sd(h$hrs1), 2)`**

The standard error of hours worked is **`r round(hrs1_sd / sqrt(hrs1_n), 4)`**

The z-score is **`r round(hrs1_diff / hrs1_se, 4)`**

The p-value, or probability of getting an absolute z-score of .21 or greater, is about **`r ((1 - pnorm(.20916488)) * 2)`**

I have a **8%** probability of getting a sample mean of at least **39.94** if the true population mean were **40**. Since the p-value of 8 is higher than alpha level 0.06, we fail to reject the null hypothesis. I can also use the 95% confidence interval for age of **39.37** and **40.51** to fail to reject the null hypothesis.

```{r}
#| echo: false
#| message: false
#| warning: false

ggplot(data.frame(x = c(39, 41)), aes(x = x)) +
  stat_function(fun = dnorm, 
                args = list(mean = hrs1_null, sd = hrs1_se), 
                color = "darkgreen") +
  stat_function(fun = dnorm, 
                args = list(mean = hrs1_null, sd = hrs1_se), 
                geom = "area", 
                xlim = c(39, 41), 
                fill = "lightblue", 
                alpha = 0.6) +
  geom_vline(xintercept = hrs1_xbar, 
             linetype = "dashed", 
             color = "black") +
  labs(x = "Mean Hours Worked", 
       y = "Density")
```


